# Basic Checking Functions
## Cpp implementation
```{r}
sourceCpp("Notebooks/Utils/checking_functions.cpp")
```

## R implementation 
### Sets

```{r}
trace.env$check_set <- function(agent_set, agents_in_room){
  if("not" %in% attr(agent_set, "modifiers")){ 
    if(is.null(agent_set) || length(agent_set) <= 0)
      return(TRUE)
    res <- !any(agent_set %in% agents_in_room)
    prefix <- trace.env$OP$negation
  }
  else{
    if(is.null(agent_set) || length(agent_set) > length(agents_in_room))
      return(FALSE)
    prefix <- ""
    res <- all(agent_set %in% agents_in_room)
  }
  # cat("\nCheck Sets: ", prefix,paste0(agent_set, collapse=","), "\t", paste0(agents_in_room, collapse=","), "\t", paste0(res, collapse=","))
  return(res)
}

trace.env$check_sets <- function(agent_set_list, agents_in_room){
  if(is.null(agent_set_list) || length(agent_set_list) <= 0){
    return(TRUE)
  }
  sets_res <- unlist(lapply(agent_set_list, check_set, agents_in_room))
  return(all(sets_res == TRUE))
}
```

### Groups

```{r}
trace.env$check_group <- function(agent_group, agents_in_room){
  if(is.null(agent_group))
    return(TRUE)
  return(any(agent_group %in% agents_in_room))
}

trace.env$check_groups <- function(agent_group_list, agents_in_room){
  if(is.null(agent_group_list) || length(agent_group_list) <= 0){
    return(TRUE)
  }
  groups_res <- unlist(lapply(agent_group_list, check_group, agents_in_room))
  return(all(groups_res == TRUE))
}
```

### Range

```{r}
trace.env$check_range <- function(agentsList, min=-Inf, max=Inf){
  total <- length(agentsList)
  if(is.null(min)){
    min = -Inf
  }
  
  if(is.null(max)){
    max = Inf
  }
  
  return(min <= total & total <= max)
}
```


## Vertical and Horizontal checks

### Indivdual Cells

Given the program data, we want to now fetch all the data related to the
info we have:

1.  Apply the size range (No point in checking the rest if the cell is
    outside of the range)

2.  Apply the sets selection (if one of the agents in the set is not
    there we can stop here)

3.  Apply the group selection

We need a function to find the unique ids in a room (which will just be
the cell if we are NOT using strings). We then need a way to count these
ids in case we have a range to check.

```{r}
## Get unique cells in a room
trace.env$unique_ids <- function(room_data){
  if(is.null(room_data)) return(c())
  if(is.list(room_data) || is.vector(room_data)){
    return(unlist(room_data))
  }
  separator <- ","
  return(unique(unlist(strsplit(unique(room_data), separator))))
}

## Get the total number of agents in the room 
trace.env$total_ids <- function(room_data){
  ids_list <- unique_ids(room_data)
  total <- length(ids_list)
  return(total)
}

```

The function `check_cell(room_data, set, min, max)` will return `TRUE`
only if the cell meets all the criteria so only if:

-   The number of agents in the room is between the minimum and the
    maximum range

-   ANY agents in the set are included in the room data

    -   Negative: Result for the cell should be NEGATED, since the check
        is done done on a positive presence basis. The agent should NOT
        be in the room

-   ALL of the agents in the group are included in the room data We can
    then apply this to the whole room's column:

```{r}
trace.env$check_cell <- function(room_data, sets, groups, range, debug=FALSE){
  in_range <- TRUE
  all_in_room <- TRUE
  at_least_one <- TRUE
  
  ids_list <- unique_ids(room_data)
  if(length(range) > 0) in_range <- check_range(ids_list, range[[1]], range[[2]])
  if(length(sets) > 0) all_in_room <- check_sets(sets, ids_list)
  if(length(groups) > 0) at_least_one <- check_groups(groups, ids_list)
  if(debug){
    cat("\nIn room: ", paste0(ids_list, collapse=","), sep="")
    cat("\nSets: ", paste0(sets, collapse=","), "\t", all_in_room, sep="")
    cat("\nGroups: ", paste0(groups, collapse=","), "\t", at_least_one, sep="")
    cat("\nRange: ", paste0(range, collapse=","), "\t", in_range, sep="")
    cat("\n")
  }
  return(in_range & all_in_room & at_least_one)
}
```
