
# Generate Synthetic dataset
I'm fairly proud of the dataset generation, it works like this:
1. Generate a regular time series
This is done by creating a sequence of times in a regular interval using `seq()`, but first we need to convert the interval to seconds.
if no `startTime` is provided the current datetime will be used. If no `endTime` is provided it is automatically calculated according to the requested rows (by simply multiplying `interval_seconds * (n-1)`)

2. For each room and for each agent:
3. Each agent starts with a column vector all set to `TRUE`, what that means is that the agent can be placed in ANY `timeSlot` at the moment, they are all available
4. Generate a random normal distribution and run it through an `ifelse`, for each value in the generated distribution check if it's over the threshold, if it is just return `TRUE` otherwise return `FALSE`
5. (if `concurrency=FALSE`Now we have a new column vector of bits. We now need to go through both column vectors in parallel and check whether they are both `TRUE` if they are it means that the slot was available but now it should not be available anymore and so we should set this to `FALSE`.
6. Finally, we need to run the resulting column vector through another `ifelse` which will return the agent name (to be later added to the vector of agents in the cell) or NA (which can easily be filtered out at the end with an `apply` over each row)
7. Now, if the room column is currently empty we need to initialize it so we'll just set the room column as the result from the `ifelse` at the previous step. Otherwise we need to combine each cell to obtain the new vector to keep in the cell, for this I'm using `Map()` which iterate through two lists in parallel. So I just need to pass the current room column and the new one with the agent name in the slots it should be and then run a function to append the new value to the one in the pre-existing vector and return that.
8. Finally, we just need to run an `apply` on each row which will check every column's cell and clean the vector in the cell to remove the `NA`s
9. To make sure the rooms at the end of the list don't end up semi-empty or completely empty I'm applying some descending threshold
- I'm generating a random distribution with the same parameters and setting the threshold its 0.8 quantile (after some experimenting)
- At each iteration I'm reducing the threshold by 20% so that the next rooms are more likely to be picked

```{r}
library(lubridate)
library(ambient) # noise
library(ggridges)
library(ggplot2)
library(data.table)
```

```{r}
generate_regular_time_series <- function(startTime=NULL, endTime=NULL, interval="00:00:10", n=100){
  if(is.null(startTime)) startTime <- Sys.time()
  time_interval <- hms::parse_hms(interval)
  seconds_interval <- (hour(time_interval) *60*60) + (minute(time_interval) * 60) + second(time_interval)
  if(is.null(endTime)) endTime <- startTime+(seconds_interval*(n-1))
  time_series_dt <- data.table(time=seq(from=startTime, to=endTime, by=seconds_interval))
  time_series_dt <- time_series_dt[, timeSlotID := .I][, c("timeSlotID", "time")]
  return(time_series_dt)
}

generate_noise_matrix <- function(nrows=100, ncols=2, frequency=0.01, octaves=3, lacunarity=2, gain=0.5, thresholds=NULL, fractal="fbm"){
  library(ambient)
  noise <- noise_perlin(c(nrows, ncols), frequency=frequency, octaves=octaves, lacunarity=lacunarity, gain=gain, fractal=fractal,pertubation="fractal")
  noise <- apply(noise, 2, function(x) abs(x* 1000))
  if(is.null(thresholds)){
  }
  # noise <- apply(noise, 2, function(x) x > threshold)
  noise <- data.table(noise)
  colnames(noise) <- paste0("r",seq(1,ncols))
  return(noise)
}

apply_thresholds <- function(noise, bins){
  thresholds_binning <- function(val, thresholds){
    n <- length(thresholds)
    prevT <- thresholds[1]
    for(i in 2:n){
      t <- thresholds[i]
      if(val >= prevT && val < t){
        val <- i-1
      }
      prevT <- t
    }
    return(val)
  }
  thresholds <- c()
  q <- 0
  bins <- bins
  for(i in 1:(bins-1)){
    thresholds <- c(thresholds, quantile(noise, (1/bins)*i)[[1]])
  }
  thresholds <- c(-Inf, thresholds, Inf)
  print(thresholds)
  # thresholds <- c(thresholds, Inf)
  return(unlist(lapply(noise, thresholds_binning, thresholds)))
}

plot_noise <- function(noiseDT, col=1){
  colName <- colnames(noiseDT)[[col]]
  return(ggplot(noiseDT, aes_string(x=colName)) + geom_density())
}

plot_noise_matrix <- function(noiseDT){
  rows <- nrow(noiseDT)
  cols <- ncol(noiseDT)
  noiseDT <- melt(noiseDT, variable.name="room", value.name="value")
  plot <- ggplot(noiseDT, aes(x=value, y=room, fill=room))
  for(c in 1:cols){
    plot <- plot + geom_density_ridges(alpha=0.6)
  }
  plot <- plot +
    theme_ridges() + 
    theme(legend.position = "none")
  return(plot)
}


generate_repeated_bools <- function(n=100, rep_ratio=0.4){
    # Set the total number of elements in the list
  total_elements <- n
  # Set the ratio of repeated elements
  repeated_ratio <- rep_ratio
  # Calculate the number of repeated elements
  num_repeated <- floor(total_elements * repeated_ratio)
  
  # Generate the repeated elements
  repeated_bools <- rep(TRUE, num_repeated)
  
  # Generate the non-repeated elements
  non_repeated_bools <- map(1:(total_elements - num_repeated), ~{
    bool <- sample(c(TRUE, FALSE), 1)
    while (bool %in% repeated_bools) {
      bool <- sample(c(TRUE, FALSE), 1)
    }
    bool
  })
  
  # Combine the repeated and non-repeated elements
  bool_list <- c(repeated_bools, non_repeated_bools)
  
  # Shuffle the list to ensure randomness
  bool_list <- sample(bool_list)
  
  # Print the final list
  return(unlist(bool_list))
}


combine_vector_columns <- function(colA, colB){
  combine_agents <- function(x, y){
    return(append(x, y))
  }
  return(Map(combine_agents, colA, colB))
}

generate_trace <- function(nrooms=5, nagents=10, ntimes=100, interval="00:00:10", concurrency=FALSE, distribution="normal", threshold_step=0.2, start_threshold=NULL, test_sort=FALSE, reps_ratio=0.3){
  cat("\nGenerating Trace with", nagents, "agents,", nrooms, "rooms and ", ntimes, "timeslots\n")
  update_available_slots <- function(is_slot_available, should_place){
    if(!should_place) return(is_slot_available)  # If we are not placing the agent here there's nothing to update
    if(!is_slot_available) return(FALSE) # if we are placing the agent but the slot is unavailable, it should keep being unavailable
    # if we are placing the agent here AND the slot is available, it should now become unavailable
    return(FALSE)
  }
  check_concurrency <- function(is_slot_available, should_place){
    if(!should_place) return(FALSE)
    # If we want to place it in this slot but the slot is unavailable we should return FALSE
    if(!is_slot_available) return(FALSE)
    # if we want to place it here and the slot is available we should return TRUE
    return(TRUE)
  }
  
  middle_sort <- function(dist){
      dist <- sort(dist)
      r_len <- length(dist)
      r_half <- ceiling(r_len/2)
      dist <- append(dist[1:r_half], sort(dist[(r_half+1):r_len], decreasing=TRUE))
      return(dist)
  }
  
  threshold <- start_threshold
  if(distribution=="normal"){
    mean <- 100
    sd <- 50
    start_quantile <- 0.85
    # generate a random distribution with the given parameters and get a 0.95 percentile
    if(is.null(threshold)){
      threshold <- quantile(rnorm(ntimes, mean = mean, sd = sd), start_quantile)
    }
  }
  else if(distribution == "uniform"){
    min <- 0
    max <- 100
    if(is.null(threshold)){
      threshold <- (min+max)/2
    }
  }
  else if(distribution == "chi"){
    df <- ntimes*0.3
    ncp <- ntimes*0.5
    if(is.null(threshold)){
      threshold <- ntimes*0.8
    }
  }
  else if(distribution == "perlin"){
    if(is.null(threshold)){
      threshold <- 250
    }
  }
  else {
    s_size <- ntimes*(1-reps_ratio)
    if(is.null(threshold)){
      threshold <- s_size*0.7
      # threshold <- ntimes*0.5
    }
  }
  ts <- generate_regular_time_series(interval=interval, n=ntimes)
  room_cols <- list()
  agents_list <- list()
  for(r in 1:nrooms){
    rCol <- paste0("r",r)
    room_cols <- append(room_cols, rCol)
    for(a in 1:nagents){
      aName <- paste0("a", a)
      if(!(aName %in% names(agents_list))){
        agents_list[[aName]] <- rep(TRUE, ntimes) # All slots are initially available
      }
      if(distribution == "normal"){
        random_slots <- rnorm(ntimes, mean = mean, sd = sd)
        random_slots <- random_slots > threshold
      }
      else if(distribution == "uniform"){
        random_slots <- runif(n=ntimes, min=min, max=max)
        random_slots <- random_slots > threshold
        if(test_sort){
          random_slots <- middle_sort(random_slots)
        }
      }
      else if(distribution == "chi"){
        random_slots <- rchisq(ntimes, df=df, ncp=ncp)
        random_slots <- random_slots > threshold
        if(test_sort){
          random_slots <- middle_sort(random_slots)
        }
      }
      else if(distribution == "perlin"){
        noise_freq <- runif(1) * 0.3
        oct <- as.integer(runif(1) * 4)
        lac <- runif(1) * 0.4
        gain <- runif(1) * 0.6
        random_slots <- generate_noise_matrix(nrows=ntimes, frequency = noise_freq, octaves = oct, lacunarity = lac, gain=gain, fractal="rigid-multi")[[1]]
        random_slots <- random_slots > threshold
      }
      else{
        # random_slots <- sample(c(1:s_size, sample(1:s_size, ntimes-s_size, replace = TRUE)))
        random_slots <- generate_repeated_bools(ntimes, reps_ratio)
      }
    
      random_slots_cleaned <- random_slots
      if(!concurrency){
        random_slots_cleaned <- Map(check_concurrency, agents_list[[aName]], random_slots)
        agents_list[[aName]] <- Map(update_available_slots, agents_list[[aName]], random_slots)
      }
      aData <- ifelse(random_slots_cleaned, aName, NA)
      if(rCol %in% colnames(ts)){
        combined <- combine_vector_columns(ts[[rCol]], aData)
        # cat("\nts: ", nrow(ts), "\taData: ", length(aData), "\tCombined: ", length(combined), "\n")
        ts[[rCol]] <- combined
      }
      else{
        # cat("\nts: ", nrow(ts), "\taData: ", length(aData), "\taData ", length(aData), "\n")
        ts[[rCol]] <- aData
      }
    }
    ts[[rCol]] <- lapply(ts[[rCol]], function(x) x[!is.na(x)])
    threshold <- threshold*(1- threshold_step)
    if(r == nrooms-1){
      threshold <- 0 # just fill the last room anyway
    }
  }
  return(ts)
}

```
```{r eval=FALSE}
# Sorting test
randomNorm <- sort(rnorm(100, 100, 50))

```
## Normal Distribution
```{r eval=FALSE}
randNormDist <- generate_trace(distribution = "normal", nrooms=3, start_threshold = 125, threshold_step = 0.15)
View(randNormDist)
invisible(get_agent_journey(randNormDist, "a1"))
```
## Uniform Distribution
```{r eval=FALSE}
randUniDist <- generate_trace(distribution = "uniform", threshold_step = 0.4, start_threshold = 70, nrooms=3)
View(randUniDist)
invisible(get_agent_journey(randUniDist, "a1"))
```
## CHI Distribution
```{r eval=FALSE}
randChiDist <- generate_trace(distribution = "chi", start_threshold = 85, threshold_step = 0.1, nrooms=3)
View(randChiDist)
invisible(get_agent_journey(randChiDist, "a1"))
```
## Sampling Distribution
```{r eval=FALSE}
randSampleDist <- generate_trace(distribution = "sample",reps_ratio = 0.35, nrooms=3)
View(randSampleDist)
invisible(get_agent_journey(randSampleDist, "a1"))
```

# Chat GPT test
```{r eval=FALSE}
# Load required libraries
library(dplyr)
library(lubridate)

# Set seed for reproducibility
set.seed(123)

# Define the number of rows and columns
num_rows <- 100
num_cols <- 3

# Generate a dataframe with a date column
df <- data.frame(date = seq(as.Date("2022-01-01"), by = "day", length.out = num_rows))
agents_list <- c("a1", "a2", "a3", "a4", "a5")
n <- length(agents_list)

# Generate random data for the additional columns
for (i in 1:num_cols) {
  col_name <- paste0("r", i)
  df[[col_name]] <- replicate(num_rows, {
    vec <- sample(agents_list, sample(1:n, 1), replace = FALSE)
    # cat("Start while")
    while (any(sapply(df[[col_name]], function(x) any(vec %in% unlist(x[-1]))))) {
      vec <- sample(agents_list, sample(1:n, 1), replace = FALSE)
    }
    # cat("-End while, ")
    vec
  })
}


# View the first few rows of the dataframe
head(df)

df <- data.table(df)
df$timeSlotID <- seq(1:nrow(df))
# colnames(df) <- c("date", "r1" ,      "r2"   ,    "r3" ,      "timeSlotID")
invisible(get_agent_journey(df, "a1"))
```


# Best Result so far
This is the best result so far where I get some decent repetitions:
```{r eval=FALSE}
randTrace <- generate_trace(distribution = "perlin", start_threshold=800, ntimes=1000)
View(randTrace)
invisible(get_agent_journey(randTrace, "a3"))
```
