# Raw RSSI export to regular time series

You'll need these files: 1. `2019Analysis/functions.R`: contains all the filtering functions 2. `export_2019-10-07_14_52_23.csv`: Raw RSSI dataset 3. `Tag_List.csv`: List of tags and their config info 4. `September_2019_Deployment.csv`: List of tags assignment 5. `deployed_devices.csv"`: List of deployed scanner devices (e.g Android phones)

```{r}
library(utils)
library(foreach) 
library(doParallel)
library(stringr)    # working with strings
library(lubridate)  # working with dates 
library(readr)
library(dplyr)
library(data.table)
library(png)
library(dlm)
numCores <- (detectCores()-1)
```

```{r}
sep2019.filename <<- "export_2019-10-07_14_52_23.csv" #Selecting the trip manually
sep2019.file <<- file(file.path("./data/original/exports/history/", sep2019.filename), 'r')  # connection for reading
sep2019.firstLine <<- readLines(sep2019.file, n = 1)
close(sep2019.file)
sep2019.firstLine
```

Read raw data RSSI export:

```{r}
sep2019 <<- fread(nThread = numCores ,file.path("./data/original/exports/history/", sep2019.filename), sep=",", colClasses=c("imei"="character", "rssi"="numeric", "tx_power"="numeric", "battery"="numeric","timestamp_scan"="numeric", "minor"="numeric"), na.strings="1c58221bc-fa61-46e2-a151-d2c4710527d0", skip = 1)
sep2019[, tx_power := NULL] #Saving memory
```

```{r}
assignID <- function(idData, dataset){
  rssiIDAssign <- function(counter,x, y){
    cat(paste(counter, x$Tag, x$From, x$To, x$ID, " (RSSI Data) "))
    y <- y[minor == x$Tag & is.na(ID) & timestamp %between% c(x$From, x$To), ID := x$ID]
    cat(" -- Done \n")
  }
  tracesIDAssign <- function(counter,x, y){
    cat(paste(counter, x$Tag, x$From, x$To, x$ID, " (Traces Data) "))
    y <- y[minor == x$Tag & is.na(ID) & (startTime %between% c(x$From, x$To) | endTime %between% c(x$From, x$To)), ID := x$ID]
    cat(" -- Done \n")
  }
  regularTracesIDAssign <- function(counter,x, y){
    cat(paste(counter, x$Tag, x$From, x$To, x$ID, " (Something Data) "))
    y <- y[minor == x$Tag & is.na(ID) & (startTime %between% c(x$From, x$To)), ID := x$ID]
    cat(" -- Done \n")
  }
  dataset <- dataset[, ID := NULL]
  dataset <- dataset[, ID := as.character()]
  rows <- nrow(idData)
  
  if("timestamp" %in% colnames(dataset)){
    cat("\nAssigning ID to Raw RSSI dataset \n")
    idData[, rssiIDAssign(paste0(.I,"/",rows), .SD, dataset), by = 1:rows]
  }
  else if("startTime" %in% colnames(dataset) && "endTime" %in% colnames(dataset)){
    cat("\nAssigning ID to traces dataset \n")
    idData[, tracesIDAssign(paste0(.I,"/",rows), .SD, dataset), by = 1:rows]
  }
  else if("startTime" %in% colnames(dataset)){
    cat("\nAssigning ID to regularised traces dataset \n")
    idData[, regularTracesIDAssign(paste0(.I,"/",rows), .SD, dataset), by = 1:rows]
  }
  
  return(dataset)
}
```

Get beacons metadata:

```{r}
gSheet.beacons <<- fread(nThread = numCores, "./data/original/Tag_List.csv")
gSheet.beacons <<- gSheet.beacons[, `:=`(TagID=as.numeric(TagID), AdvRate=as.numeric(AdvRate), TxPower=as.numeric(TxPower))]
sep2019.testBatch.beacons <- gSheet.beacons[stringr::str_detect(Position, 'Test Batch')] #Getting the test minors

gSheet.deployment <<- fread(nThread = numCores, "./data/original/September_2019_Deployment.csv")
gSheet.deployment$From <- as.POSIXct(gSheet.deployment$From, tz="Australia/Melbourne", origin = "1970-01-01",  format = "%d/%m/%Y %H:%M", optional=TRUE)
gSheet.deployment$To <- as.POSIXct(gSheet.deployment$To, tz="Australia/Melbourne", origin = "1970-01-01",  format = "%d/%m/%Y %H:%M", optional=TRUE)
gSheet.deployment[is.na(From), From := sep2019.timeRange$minTime][is.na(To), To := sep2019.timeRange$maxTime]
gSheet.deployment[, Tag := as.numeric(Tag)]
gSheet.deployment[,Total := (To-From)] #If beacon 1 is assigned in the ranges 1-6 and 2-3
gSheet.deployment <- gSheet.deployment[order(Total)] #If beacon 1 is assigned in the ranges 1-6 and 2-3

sep2019.nurse.beacons <- gSheet.deployment[stringr::str_detect(Role2, 'urse')] #Nurses
sep2019.patients.beacons <- gSheet.deployment[stringr::str_detect(Role2, 'atient')] #Patients
sep2019.staff.beacons <- gSheet.deployment[Role2 %in% c("PSA", "Theatre Tech", "Head Theatre Tech")] #Patients
```

Get devices' data:

```{r}
devices <<- fread(nThread = numCores, "./data/original/deployed_devices.csv", colClasses = c("id"="character", "nroom"="integer"))
devices[, `:=`(room=paste(nroom, roomShort, sep="-"), roomChar=roomCharLong)]
devicesLetters <- c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J", "K", "L", "M", "N", "O", "P", "Q", "R", "S", "T", "U", "V", "W", "X", "Y", "Z")
devices[, roomChar := devicesLetters[.I] ]
```

Assign ID from deployment sheet

```{r}
sep2019 <- assignID(gSheet.deployment, sep2019)
sep2019.filtered <- sep2019[!is.na(ID)]
```

Apply the median filter to smooth traces:

```{r}
sep2019.smoothed15 <- filterMedianMaximumRssiDT(dt=sep2019.filtered, timeWindow = 15, selectMaximum = TRUE, intervalPerMinor = FALSE, preKalman=FALSE, useRssiPacketScoring = TRUE)
sep2019.smoothed15 <- merge(x=sep2019.smoothed15, y=devices[,c("id", "room")], by.x="imei", by.y="id")
assignID(gSheet.deployment, sep2019.smoothed15)
sep2019.smoothed15 <- sep2019.smoothed15[order(from)]
```

Generate traces:

```{r}
sep2019.traces15 <- generateTrips(filteredRecords = sep2019.smoothed15, timeBeforeLosingTrack = 300, rssiThreshold = -99, connectSameRooms = TRUE, connectDifferentRooms = FALSE, tripPointsThreshold = 2, usePointsThreshold = TRUE, nThread = 1)

```

Regularise time series:

```{r}
timeRegularSeries <- data.frame(startTime=seq(min(sep2019.traces15$startTime), max(sep2019.traces15$endTime), by=10))
timeRegularSeries$endTime <- append(timeRegularSeries$startTime[-1], timeRegularSeries$startTime[length(timeRegularSeries$startTime)]+10)
sep2019.traces15.regular <- generateRegularTimeSeries(keepLastRoomOnLost = FALSE, data=sep2019.traces15, timeRegularSeries=timeRegularSeries, devices = devices, minors=unique(sep2019.traces15$minor), timeBreaks=10)
```

Make sure column types are correct:

```{r}
tz(sep2019.traces15$startTime) <- "Australia/Melbourne"
tz(sep2019.traces15$endTime) <- "Australia/Melbourne"
setDT(sep2019.traces15)
assignID(gSheet.deployment, sep2019.traces15)
sep2019.traces15 <- sep2019.traces15[order(startTime)]
sep2019.traces15 <- merge(sep2019.traces15, devices[,c("nroom", "room")],by.x="startRoom", by.y="nroom")
sep2019.traces15.timeRange <- sep2019.traces15[, .(min=min(startTime), max=max(endTime)), by=room]
sep2019.traces15.timeRange[, diff:=as.numeric(difftime(max,min, units="secs"))]
```
