---
title: "Tool"
output: html_notebook
editor_options: 
  markdown: 
    wrap: 72
---

# First step: Read the dataset and transform it into a data.table

```{r}
library(dplyr)
library(data.table)
library(rmdHelpers)
library(utils)
library(foreach) 
library(doParallel)
library(stringr)    # working with strings
library(lubridate)  # working with dates 
library(readr)
library(dlm)
numCores <- (detectCores()-1)
```

```{r}
setwd("/Users/marinig/Documents/GitHub/localization_toolkit/R_Tool")
load("data/sep2019.traces15.regular.RData")
load("data/devices.RData")
sourceCpp('tool_functions.cpp')
```

Set up the new dataset: (Remember to use `<<-` to assign new.tripsDT in
the global environment so that the `combine_datasets` function can
access and manipulate it)

Populate the new dataset from the regularised trips:

```{r}
asString <- FALSE

beaconsData <- sep2019.traces15.regular$data
timeSeries <- sep2019.traces15.regular$timeSeries[["startTime"]]
roomNames <<- sort(devices[nroom > 0]$nroom)
beaconsInDataset <- unlist(lapply(beaconsData, function(x) x$minor[1]))

new.tripsDT <<- build_trace_dataset(timeSeries, roomNames)
# Combines each individual beacon's dataset into one large dataset, the result is saved into `new.tripsDt`
invisible(lapply(beaconsData, combine_datasets, new.tripsDT, beaconsList = beaconsInDataset, asString=asString)) # asString=FALSE to have each cell as a vector instead of a string

## Clean the strings in case asString == TRUE
if(asString){
  for (j in colnames(new.tripsDT)[-1]) set(new.tripsDT, j = j, value = substring(new.tripsDT[[j]], 2))
}

# adding a new column called timeSlotID but I want it to be in front, so I'm saving the previous columns order and then adding it in front when reordering
prevColNames <- colnames(new.tripsDT)
new.tripsDT[, timeSlotID := .I]
setcolorder(new.tripsDT, c("timeSlotID", prevColNames))
```

Just testing that the cells are ACTUALLY vectors

```{r}
new.tripsDT[,"14"][[1]][[1]]
```

# Second step: Query testing

## Run simple query

Let's run a simple query just to make sure the data structure makes
sense:

```{r}
vectorized_contains <- function(vector, element){
  return(element %in% vector)
}
sample  <- new.tripsDT[1:100000]
cat(sample[['33']][[1]]) # Test
# Right away we can see that beacon 334 moves from room 14 to room 24 at time slots 20 and 22, let's try to capture that
sampleFromRoom <- "14"
sampleToRoom <- "24"
sampleBeacon <- "334"
sampleResultFrom <- mapply(vectorized_contains, sample[[sampleFromRoom]], sampleBeacon)
sampleResultTo <- mapply(vectorized_contains, sample[[sampleToRoom]], sampleBeacon)
sampleResult <- sample[sampleResultFrom | sampleResultTo]
sampleResult

```

## Find jumps

Let's look for the jumps in time. We calculated the difference between
the timeSlotID of the current row and the previous one by using shift
with type='lag'

```{r}
## Let's look for the jumps
totalJumps <- sampleResult[ , diff := timeSlotID - shift(timeSlotID, 1L, type="lag")]
totalJumps[diff > 1] # This should return 0 as new.tripsDT contains ALL the time slots.
```

Now let's have a look at the jumps in the "from" room. We can also
search for the timeSlotID before the jump by using the lookahead shift
`shift(colName, 1L, type="lead")` We can see that there is one jump,
beacon 334 appears in room

```{r}
cols <- c('timeSlotID', 'time', 'diff', sampleFromRoom, sampleToRoom)
sampleTotalFromJumps <- sample[sampleResultFrom, diff := timeSlotID - shift(timeSlotID, 1L, type="lag")]
sampleTotalFromJumps[diff > 1 | shift(diff, 1L, type="lead") > 1, ..cols]
```

```{r}
cols <- c('timeSlotID', 'time', 'diff', sampleFromRoom, sampleToRoom)
sampleTotalToJumps <- sample[sampleResultTo , diff := timeSlotID - shift(timeSlotID, 1L, type="lag")]
sampleTotalToJumps[diff > 1 | shift(diff, 1L, type="lead") > 1, ..cols]
```

Finally, let's have a look at all the jumps the beacon made between the
two rooms. We can see that beacon 334 moved between room 14 and 24 a
total of 3 times:

1.  Between tID 20 and 22

2.  Between tID 8636 and 8641

3.  Between tID 8642 and 8644

```{r}
cols <- c('timeSlotID', 'time', 'diff', sampleFromRoom, sampleToRoom)
sampleTotalJumps <- sampleResult[ , diff := timeSlotID - shift(timeSlotID, 1L, type="lag")]
sampleTotalJumps[diff > 1 | shift(diff, 1L, type="lead") > 1, ..cols]
```

## Extract jumps data

Let's investigate further. Let's extract those jumps from the full
data.table and see what's happening

```{r}
jumpRowFrom <- 20
jumpRowTo <- 22
padding <- 2
jump1 <- new.tripsDT[timeSlotID >= jumpRowFrom-padding & timeSlotID <= jumpRowTo+padding]
jump1
```

### Filter columns

Let's filter the columns to only get the time-related ones. This will
only return TRUE for the three columns we listed

```{r}
colnames(jump1) %in% c('timeSlotID', 'time', 'diff')
```

This will return True only for the columns where at least one row
contains the beacon's ID

```{r}
sapply(jump1, function(x) any(x %like% sampleBeacon))
```

Now we just need to OR them and get TRUE for both the 3 columns we want
and the ones where the beacon appears

```{r}
jump1 <- jump1[, colnames(jump1) %in% c('timeSlotID', 'time', 'diff') | sapply(jump1, function(x) any(x %like% sampleBeacon)), with=F]
print(colnames(jump1))
jump1
```

## Check the other jumps

Let's do the same for the other two jumps:

```{r}
jumpRowFrom <- 8636
jumpRowTo <- 8641
padding <- 2
jump2 <- new.tripsDT[timeSlotID >= jumpRowFrom-padding & timeSlotID <= jumpRowTo+padding]
jump2 <- jump2[, colnames(jump2) %in% c('timeSlotID', 'time', 'diff') | sapply(jump2, function(x) any(x %like% sampleBeacon)), with=F]
print(colnames(jump2))
jump2
```

```{r}
jumpRowFrom <- 8642
jumpRowTo <- 8644
padding <- 2
jump3 <- new.tripsDT[timeSlotID >= jumpRowFrom-padding & timeSlotID <= jumpRowTo+padding]
jump3 <- jump3[, colnames(jump3) %in% c('timeSlotID', 'time', 'diff') | sapply(jump3, function(x) any(x %like% sampleBeacon)), with=F]
print(colnames(jump3))
jump3
```

## Agents Selection Testing

We can run some testing for the previous example. We know beacon 334
moved from room 14 at t=20 and came back at some point around t=8641

```{r}
checkedCol <- check_room_series(sample[["14"]], "334")
cat("How many time slots was 334 in the room 14?: ", length(which(checkedCol == TRUE)), "\n\n")
cat("How many time slots was 334 NOT in the room 14?: ", length(which(checkedCol == FALSE)), "\n\n")
```

Let's try with some range testing

```{r}
total_in_room <- lapply(sample[["14"]], total_ids)
sample_min5 <- which(total_in_room > 5)
sample_max1 <- which(total_in_room < 1)
cat("\n", length(sample_min5), "Time slots with > 5 people:\n", paste0(sample_min5))
cat("\n\n", length(sample_max1), "Time slots with < 1 people (empty):\n", paste0(sample_max1))
```

We can also see what agents 334 was with at any point in time, a list of
their encounters:

```{r}
unique_ids_in_room <- lapply(sample[["14"]], unique_ids)
unique_combinations <- unique(unique_ids_in_room) # Getting the unique combinations of IDs
which(unique_combinations %in% "334") # This does not work as it will only get the combination which is exactly "334"
# We need to run lapply to check each list in the list
idx_with_334 <- which(lapply(unique_combinations, function(x) "334" %in% x) == TRUE)
invisible(lapply(unique_combinations[idx_with_334], function(x) cat(x, "\n\n")))
```

Finally, we can now try to combine all together to get ranges and sets.
Let's get all the time slots where 334 was in room 14 with at least 0
and at most 1 person. So they should be alone

```{r}
cols <- c('timeSlotID', 'time', 'diff', "14")
checkedCol334Alone <- check_room_series(sample[["14"]], "334", 0, 1)
sample[(checkedCol334Alone), ..cols]
```

PERFECT!! Now one more time but with at most 2 people:

```{r}
cols <- c('timeSlotID', 'time', 'diff', "14")
checkedCol334 <- check_room_series(sample[["14"]], "334", 0, 2)
sample[checkedCol334, ..cols]
```

Technically that should mean that if we take the previous indexes, we
should be able to get the times where 334 was alone (again):

```{r}
intersection <- unlist(checkedCol334) & unlist(checkedCol334Alone)
sample[intersection, ..cols]
cat(length(checkedCol334Alone), " vs ", length(intersection))
```


# Synthetic Tests
```{r}
generate_
```


